The effects of drought on land cover change and vegetation productivity in continental Chile

### 2024-06-25 article/article_first_version/paper_drought_Chile2023.qmd
actualizar repositorio0+

To analyze the land cover change, we use the IGBP scheme from the MCD12Q1 Collection 6.1 from MODIS. This product has been previously used for studies of drought and land cover in Chile [@Fuentes2021;@Zambrano2018]. We regrouped the 17 classes into ten macroclasses, as follows: classes 1-4 to forest, 5-7 to shrublands, 8-9 to savannas, 10 as grasslands, 11 as wetlands, 12 and 14 to croplands, 13 as urban, 15 as snow and ice, 16 as barren, and 17 to water bodies (Table S1). Thus, we have a land cover raster time series with the ten macroclasses for 2001 and 2023. We validate the land cover macroclasses regarding a highly detailed (30 m of spatial resolution) land cover map made for Chile by @Zhao2016 for 2013-2014. Our results showed a global accuracy of ~0.82 and a F1 score of ~0.66. Section S2 in the Supplementary Material shows the procedure for validation.

>>>>

To analyze the land cover change, we use the IGBP scheme from the MCD12Q1 Collection 6.1 from MODIS. This product has been previously used for studies of drought and land cover in Chile [@Fuentes2021;@Zambrano2018]. We regrouped the 17 classes into ten macroclasses, as follows: classes 1-4 to forest, 5-7 to shrublands, 8-9 to savannas, 10 as grasslands, 11 as wetlands, 12 and 14 to croplands, 13 as urban, 15 as snow and ice, 16 as barren, and 17 to water bodies (see Table S2). Thus, we have a land cover raster time series with the ten macroclasses for 2001 and 2023. We validate the land cover macroclasses regarding a highly detailed (30 m of spatial resolution) land cover map made for Chile by @Zhao2016 for 2013-2014. Our results showed a global accuracy of ~0.82 and a F1 score of ~0.66. Section S2 in the Supplementary Material shows the procedure for validation.

### 2024-06-25 article/article_first_version/paper_drought_Chile2023.tex
actualizar repositorio0+

\usepackage{lineno}\linenumbers \usepackage{multirow} \usepackage{lscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The effects of drought on land cover change and vegetation productivity in continental Chile},
  pdfauthor={Francisco Zambrano; Anton Vrieling; Francisco Meza; Iongel Duran-Llacer; Francisco Fernández; Alejandro Venegas-González; Nicolas Raab; Dylan Craven},
  pdfkeywords={drought, land cover change, vegetation
productivity, ecosystem},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{The effects of drought on land cover change and vegetation
productivity in continental Chile}
\author[1,2]{Francisco Zambrano%
\corref{cor1}%
}
 \ead{francisco.zambrano@umayor.cl} 
\author[3]{Anton Vrieling%
%
}
 \ead{a.vrieling@utwente.nl} 
\author[4,5,6]{Francisco Meza%
%
}

@@

\author[8,9]{Francisco Fernández%
%
}
\author[10]{Alejandro Venegas-González%
%
}

@@

\affiliation[1]{organization={Hémera Centro de Observación de la Tierra,
Facultad de Ciencias, Escuela de Ingeniería en Medio Ambiente y
Sustentabilidad, Universidad Mayor},city={Santiago},country={7500994,
Chile.}}
\affiliation[2]{organization={Observatorio de Sequía para la Agricultura
y la Biodiversidad de Chile (ODES), Universidad
Mayor},city={Santiago},country={7500994,
Chile.}}
\affiliation[3]{organization={Faculty of Geo-Information Science and
Earth, University of Twente},city={Enschede},country={The
Netherlands.}}
\affiliation[4]{organization={Facultad de Agronomía y Sistemas
Naturales, Pontificia Universidad Católica de
Chile.},city={Santiago},country={Chile.}}
\affiliation[5]{organization={Instituto para el Desarrollo Sustentable.
Pontificia Universidad Católica de
Chile},city={Santiago},country={Chile.}}
\affiliation[6]{organization={Centro Interdisciplinario de Cambio
Global, Pontificia Universidad Católica de
Chile},city={Santiago},country={Chile.}}
\affiliation[7]{organization={Hémera Centro de Observación de la Tierra,
Facultad de Ciencias, Universidad
Mayor,},city={Santiago},country={7500994,
Chile.}}
\affiliation[8]{organization={Center of Economics for Sustainable
Development (CEDES), Faculty of Economics and Government, Universidad
San
Sebastian},city={Santiago},country={Chile.}}
\affiliation[9]{organization={Center of Applied Ecology and
Sustainability
(CAPES)},city={Santiago},country={Chile.}}
\affiliation[10]{organization={Instituto de Ciencias Agroalimentarias,
Animales y Ambientales (ICA3), Universidad de O'Higgins},city={San
Fernando},country={Chile.}}
\affiliation[11]{organization={, GEMA Center for Genomics, Ecology \&
Environment, Universidad Mayor, Camino La Pirámide Huechuraba
5750},city={Santiago},country={Chile.}}
\affiliation[12]{organization={Data Observatory
Foundation},city={Santiago},country={Chile.}}

@@

\begin{keyword}
    drought \sep land cover change \sep vegetation productivity \sep 
    ecosystem
\end{keyword}
\end{frontmatter}

\captionsetup{justification=raggedright,singlelinecheck=false}

\section{Introduction}\label{introduction}

@@

\section{Study area}\label{study-area}

@@

\centering{

\includegraphics{../output/figs/map_study_con_landcover.png}

@@

\end{figure*}%

\section{Materials and Methods}\label{materials-and-methods}

\subsection{Data}\label{data}

\subsubsection{Gridded meteorological and vegetation
data}\label{gridded-meteorological-and-vegetation-data}

@@

\begin{table}

\centering{

[!ht]

\scriptsize
\centering
\begin{tabular}{p{0.1\textwidth}cp{0.3\textwidth}p{0.095\textwidth}ccc}
\hline
\multirow{1}{*}{\centering Product} & Sub-product & Variable & Spatial Resolution  & Period & Units & Short Name \\ 
\hline
\multirow{4}{*}{ERA5L} & ~ & Precipitation & \multirow{4}{*}{~0.1°} & \multirow{4}{*}{1981-2023} & mm & P \\ 
         &  & Maximum temperature & ~ & & $°C$ & $T_{max}$ \\ 
         &  & Minimum temperature & ~ & & $°C$ & $T_{min}$ \\ 
         &  & Volumetric Soil Water Content at 1m & ~ & & $m3/m3$ & SM \\ 
ERA5L* & & Atmospheric Evaporative Demand & 0.1° & 1981-2023 & mm & AED \\
        \multirow{2}{*}{MODIS} & MOD13A3.061 & Normalized Difference Vegetation Index & \multirow{2}{*}{~1 km} & 2000-2023 & ~ & NDVI \\ 
         & MCD12Q1.061 & land cover IGBP scheme & & 2001-2022 & ~ & land cover \\ 
\hline
\end{tabular}
{\raggedright *Calculated from maximum and minimum temperatures derived from ERA5L with Eq. \ref{eq-AED}. \par}

\caption{\label{tab-desEOD}Description of the satellite and reanalysis data used}

\end{table}%

@@

\subsection{Short- to long-term drought
trends}\label{short--to-long-term-drought-trends}

\subsubsection{Atmospheric Evaporative Demand
(AED)}\label{atmospheric-evaporative-demand-aed}

@@

\begin{equation}\phantomsection\label{eq-AED}{AED = 0.0023\cdot Ra\cdot (T+17.8)\cdot (T_{max}-T_{min})^{0.5}}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-Ra}{R_a = \frac{14,400}{\pi}\cdot G_{sc}\cdot d_r \left[\omega_s\cdot sin(\phi)\cdot sin(\delta)+cos(\phi)\cdot cos(\delta)\cdot sin(\omega_s) \right]}\end{equation}

@@

\subsubsection{Non-parametric calculation of drought
indices}\label{non-parametric-calculation-of-drought-indices}

@@

\begin{equation}\phantomsection\label{eq-sumvar}{A^s_i = \sum_{i=n-s-i+2}^{n-i+1} V_i\,\, \forall\, i\geq n-s+1  }\end{equation}

@@

\begin{equation}\phantomsection\label{eq-probPai}{P(A^s_i) = \frac{i-0.33}{n+0.33'}}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-DI}{DI(A^s_i) = W - \frac{C_0+C_1\cdot W + c_2 \cdot W^2}{1+d_1\cdot W +d_2\cdot W^2 +d_3\cdot W^3}}\end{equation}

@@

\subsubsection{Trend of drought indices}\label{trend-of-drought-indices}

@@

\subsection{Interaction of land cover and
drought}\label{interaction-of-land-cover-and-drought}

\subsubsection{Land cover change}\label{land-cover-change}

@@

\subsubsection{Relationship between land cover and drought
trends}\label{relationship-between-land-cover-and-drought-trends}

@@

\subsection{Drought impacts on vegetation
productivity}\label{drought-impacts-on-vegetation-productivity}

@@

\subsection{Software}\label{software}

@@

\section{Results}\label{results}

\subsection{Short- to long-term drought
trends}\label{short--to-long-term-drought-trends-1}

@@

\begin{minipage}{0.50\linewidth}

\centering{

@@

\subcaption{\label{fig-trendDI-1}SPI (Standardized Precipitation Index)}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

@@

\subcaption{\label{fig-trendDI-2}EDDI (Evaporative Demand Drought
Index)}

\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}

\centering{

@@

\subcaption{\label{fig-trendDI-3}SPEI (Standardized Precipitation
Evapotranspiration Index)}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

@@

\subcaption{\label{fig-trendDI-4}SSI (Standardized Soil Moisture Index)}

\end{minipage}%

@@

\end{figure*}%

@@

\centering{

\includegraphics{../output/figs/trend_macrozone_drought_indices.png}

@@

\end{figure}%

@@

\subsection{Interaction of land cover and
drought}\label{interaction-of-land-cover-and-drought-1}

\subsubsection{Land cover change}\label{land-cover-change-1}

@@

\centering{

\includegraphics{../output/figs/proportion_landcover_macroclass_2001-2022.png}

@@

\subsubsection{Relationship between drought indices and land cover
change}\label{relationship-between-drought-indices-and-land-cover-change}

@@

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{../output/figs/bars_relative_importance_RF.png}

@@

\centering{

\includegraphics{../output/figs/points_landcover_drought_indices_trend_and_time_scale.png}

@@

\subsection{Drought impacts on vegetation productivity within land
cover}\label{drought-impacts-on-vegetation-productivity-within-land-cover}

\subsubsection{Trends in vegetation
productivity}\label{trends-in-vegetation-productivity}

@@

\centering{

\includegraphics{../output/figs/temporal_variation_zcNDVI6_macrozonas_con_mapa.png}

@@

\subsubsection{Correlation between vegetation productivity and drought
indices}\label{correlation-between-vegetation-productivity-and-drought-indices}

@@

\centering{

\includegraphics{../output/figs/mapa_cor_selec_indices_zcNDVI6.png}

@@

\centering{

\includegraphics{../output/figs/mapa_cor_r_indices_zcNDVI6.png}

@@

\section{Discussion}\label{discussion}

\subsection{Vegetation water demand and its relation to
drought}\label{vegetation-water-demand-and-its-relation-to-drought}

@@

\subsection{Sensitivity of land cover vegetation to short- and long-term
drought}\label{sensitivity-of-land-cover-vegetation-to-short--and-long-term-drought}

@@

\subsection{Vegetation productivity and
drought}\label{vegetation-productivity-and-drought}

@@

\subsection{Drought information to aid in
adaptation}\label{drought-information-to-aid-in-adaptation}

@@

\section{Conclusion}\label{conclusion}

@@

\section{Acknowledgment}\label{acknowledgment}

The National Research and Development Agency of Chile (ANID) funded this
study through the drought emergency project FSEQ210022, Fondecyt
Iniciación N°11190360, Fondecyt Postdoctorado N°3230678, and Fondecyt
Regular N°1210526.

>>>>

\usepackage{lineno}\linenumbers \usepackage{multirow} \usepackage{lscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The effects of drought on land cover change and vegetation productivity in continental Chile},
  pdfauthor={Francisco Zambrano; Anton Vrieling; Francisco Meza; Iongel Duran-Llacer; Francisco Fernández; Alejandro Venegas-González; Nicolas Raab; Dylan Craven},
  pdfkeywords={drought, land cover change, vegetation
productivity, ecosystem},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{The effects of drought on land cover change and vegetation
productivity in continental Chile}
\author[1,2]{Francisco Zambrano%
\corref{cor1}%
}
 \ead{francisco.zambrano@umayor.cl} 
\author[3]{Anton Vrieling%
%
}

\author[4,5,6]{Francisco Meza%
%
}

@@

\author[8,9]{Francisco Fernández%
%
}

\author[10]{Alejandro Venegas-González%
%
}

@@

\affiliation[1]{organization={Hémera Centro de Observación de la Tierra,
Facultad de Ciencias, Escuela de Ingeniería en Medio Ambiente y
Sustentabilidad, Universidad Mayor},city={Santiago},country={7500994,
Chile.},countrysep={,},postcodesep={}}
\affiliation[2]{organization={Observatorio de Sequía para la Agricultura
y la Biodiversidad de Chile (ODES), Universidad
Mayor},city={Santiago},country={7500994,
Chile.},countrysep={,},postcodesep={}}
\affiliation[3]{organization={Faculty of Geo-Information Science and
Earth, University of Twente},city={Enschede},country={The
Netherlands.},countrysep={,},postcodesep={}}
\affiliation[4]{organization={Facultad de Agronomía y Sistemas
Naturales, Pontificia Universidad Católica de
Chile.},city={Santiago},country={Chile.},countrysep={,},postcodesep={}}
\affiliation[5]{organization={Instituto para el Desarrollo Sustentable.
Pontificia Universidad Católica de
Chile},city={Santiago},country={Chile.},countrysep={,},postcodesep={}}
\affiliation[6]{organization={Centro Interdisciplinario de Cambio
Global, Pontificia Universidad Católica de
Chile},city={Santiago},country={Chile.},countrysep={,},postcodesep={}}
\affiliation[7]{organization={Hémera Centro de Observación de la Tierra,
Facultad de Ciencias, Universidad
Mayor,},city={Santiago},country={7500994,
Chile.},countrysep={,},postcodesep={}}
\affiliation[8]{organization={Center of Economics for Sustainable
Development (CEDES), Faculty of Economics and Government, Universidad
San
Sebastian},city={Santiago},country={Chile.},countrysep={,},postcodesep={}}
\affiliation[9]{organization={Center of Applied Ecology and
Sustainability
(CAPES)},city={Santiago},country={Chile.},countrysep={,},postcodesep={}}
\affiliation[10]{organization={Instituto de Ciencias Agroalimentarias,
Animales y Ambientales (ICA3), Universidad de O'Higgins},city={San
Fernando},country={Chile.},countrysep={,},postcodesep={}}
\affiliation[11]{organization={},,postcodesep={}}
\affiliation[12]{organization={, GEMA Center for Genomics, Ecology \&
Environment, Universidad Mayor, Camino La Pirámide Huechuraba
5750},city={Santiago},country={Chile.},countrysep={,},postcodesep={}}
\affiliation[13]{organization={Data Observatory
Foundation},city={Santiago},country={Chile.},countrysep={,},postcodesep={}}

@@

\begin{keyword}
    drought \sep land cover change \sep vegetation productivity \sep 
    ecosystem
\end{keyword}
\end{frontmatter}
    \captionsetup{justification=raggedright,singlelinecheck=false}

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, sharp corners, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\hypertarget{study-area}{%
\section{Study area}\label{study-area}}

@@

{\centering \includegraphics{../output/figs/map_study_con_landcover.png}

@@

\end{figure*}

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

\hypertarget{gridded-meteorological-and-vegetation-data}{%
\subsubsection{Gridded meteorological and vegetation
data}\label{gridded-meteorological-and-vegetation-data}}

@@

\begin{table}[!ht]
\caption{Description of the satellite and reanalysis data used}
\label{tab-desEOD}
\scriptsize
\centering
\begin{tabular}{p{0.1\textwidth}cp{0.3\textwidth}p{0.095\textwidth}ccc}
\hline
\multirow{1}{*}{\centering Product} & Sub-product & Variable & Spatial Resolution  & Period & Units & Short Name \\ 
\hline
\multirow{4}{*}{ERA5L} & ~ & Precipitation & \multirow{4}{*}{~0.1°} & \multirow{4}{*}{1981-2023} & mm & P \\ 
         &  & Maximum temperature & ~ & & $°C$ & $T_{max}$ \\ 
         &  & Minimum temperature & ~ & & $°C$ & $T_{min}$ \\ 
         &  & Volumetric Soil Water Content at 1m & ~ & & $m3/m3$ & SM \\ 
ERA5L* & & Atmospheric Evaporative Demand & 0.1° & 1981-2023 & mm & AED \\
        \multirow{2}{*}{MODIS} & MOD13A3.061 & Normalized Difference Vegetation Index & \multirow{2}{*}{~1 km} & 2000-2023 & ~ & NDVI \\ 
         & MCD12Q1.061 & land cover IGBP scheme & & 2001-2022 & ~ & land cover \\ 
\hline
\end{tabular}
{\raggedright *Calculated from maximum and minimum temperatures derived from ERA5L with Eq. \ref{eq-AED}. \par}
\end{table}

@@

\hypertarget{short--to-long-term-drought-trends}{%
\subsection{Short- to long-term drought
trends}\label{short--to-long-term-drought-trends}}

\hypertarget{atmospheric-evaporative-demand-aed}{%
\subsubsection{Atmospheric Evaporative Demand
(AED)}\label{atmospheric-evaporative-demand-aed}}

@@

\begin{equation}\protect\hypertarget{eq-AED}{}{AED = 0.0023\cdot Ra\cdot (T+17.8)\cdot (T_{max}-T_{min})^{0.5}}\label{eq-AED}\end{equation}

@@

\begin{equation}\protect\hypertarget{eq-Ra}{}{R_a = \frac{14,400}{\pi}\cdot G_{sc}\cdot d_r \left[\omega_s\cdot sin(\phi)\cdot sin(\delta)+cos(\phi)\cdot cos(\delta)\cdot sin(\omega_s) \right]}\label{eq-Ra}\end{equation}

@@

\hypertarget{non-parametric-calculation-of-drought-indices}{%
\subsubsection{Non-parametric calculation of drought
indices}\label{non-parametric-calculation-of-drought-indices}}

@@

\begin{equation}\protect\hypertarget{eq-sumvar}{}{A^s_i = \sum_{i=n-s-i+2}^{n-i+1} V_i\,\, \forall\, i\geq n-s+1  }\label{eq-sumvar}\end{equation}

@@

\begin{equation}\protect\hypertarget{eq-probPai}{}{P(A^s_i) = \frac{i-0.33}{n+0.33'}}\label{eq-probPai}\end{equation}

@@

\begin{equation}\protect\hypertarget{eq-DI}{}{DI(A^s_i) = W - \frac{C_0+C_1\cdot W + c_2 \cdot W^2}{1+d_1\cdot W +d_2\cdot W^2 +d_3\cdot W^3}}\label{eq-DI}\end{equation}

@@

\hypertarget{trend-of-drought-indices}{%
\subsubsection{Trend of drought
indices}\label{trend-of-drought-indices}}

@@

\hypertarget{interaction-of-land-cover-and-drought}{%
\subsection{Interaction of land cover and
drought}\label{interaction-of-land-cover-and-drought}}

\hypertarget{land-cover-change}{%
\subsubsection{Land cover change}\label{land-cover-change}}

@@

\hypertarget{relationship-between-land-cover-and-drought-trends}{%
\subsubsection{Relationship between land cover and drought
trends}\label{relationship-between-land-cover-and-drought-trends}}

@@

\hypertarget{drought-impacts-on-vegetation-productivity}{%
\subsection{Drought impacts on vegetation
productivity}\label{drought-impacts-on-vegetation-productivity}}

@@

\hypertarget{software}{%
\subsection{Software}\label{software}}

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{short--to-long-term-drought-trends-1}{%
\subsection{Short- to long-term drought
trends}\label{short--to-long-term-drought-trends-1}}

@@

\begin{figure}

\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

@@

\subcaption{\label{fig-trendDI-1}SPI (Standardized Precipitation Index)}
\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

@@

\subcaption{\label{fig-trendDI-2}EDDI (Evaporative Demand Drought
Index)}
\end{minipage}%
\newline
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

@@

\subcaption{\label{fig-trendDI-3}SPEI (Standardized Precipitation
Evapotranspiration Index)}
\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

@@

\subcaption{\label{fig-trendDI-4}SSI (Standardized Soil Moisture Index)}
\end{minipage}%

@@

\end{figure}

@@

{\centering \includegraphics{../output/figs/trend_macrozone_drought_indices.png}

@@

\end{figure}

@@

\hypertarget{interaction-of-land-cover-and-drought-1}{%
\subsection{Interaction of land cover and
drought}\label{interaction-of-land-cover-and-drought-1}}

\hypertarget{land-cover-change-1}{%
\subsubsection{Land cover change}\label{land-cover-change-1}}

@@

{\centering \includegraphics{../output/figs/proportion_landcover_macroclass_2001-2022.png}

@@

\hypertarget{relationship-between-drought-indices-and-land-cover-change}{%
\subsubsection{Relationship between drought indices and land cover
change}\label{relationship-between-drought-indices-and-land-cover-change}}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../output/figs/bars_relative_importance_RF.png}

@@

{\centering \includegraphics{../output/figs/points_landcover_drought_indices_trend_and_time_scale.png}

@@

\hypertarget{drought-impacts-on-vegetation-productivity-within-land-cover}{%
\subsection{Drought impacts on vegetation productivity within land
cover}\label{drought-impacts-on-vegetation-productivity-within-land-cover}}

\hypertarget{trends-in-vegetation-productivity}{%
\subsubsection{Trends in vegetation
productivity}\label{trends-in-vegetation-productivity}}

@@

{\centering \includegraphics{../output/figs/temporal_variation_zcNDVI6_macrozonas_con_mapa.png}

@@

\hypertarget{correlation-between-vegetation-productivity-and-drought-indices}{%
\subsubsection{Correlation between vegetation productivity and drought
indices}\label{correlation-between-vegetation-productivity-and-drought-indices}}

@@

{\centering \includegraphics{../output/figs/mapa_cor_selec_indices_zcNDVI6.png}

@@

{\centering \includegraphics{../output/figs/mapa_cor_r_indices_zcNDVI6.png}

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{vegetation-water-demand-and-its-relation-to-drought}{%
\subsection{Vegetation water demand and its relation to
drought}\label{vegetation-water-demand-and-its-relation-to-drought}}

@@

\hypertarget{sensitivity-of-land-cover-vegetation-to-short--and-long-term-drought}{%
\subsection{Sensitivity of land cover vegetation to short- and long-term
drought}\label{sensitivity-of-land-cover-vegetation-to-short--and-long-term-drought}}

@@

\hypertarget{vegetation-productivity-and-drought}{%
\subsection{Vegetation productivity and
drought}\label{vegetation-productivity-and-drought}}

@@

\hypertarget{drought-information-to-aid-in-adaptation}{%
\subsection{Drought information to aid in
adaptation}\label{drought-information-to-aid-in-adaptation}}

@@

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

@@

\hypertarget{acknowledgment}{%
\section{Acknowledgment}\label{acknowledgment}}

The National Research and Development Agency of Chile (ANID) funded this
study through the drought emergency project FSEQ210022, Fondecyt
Iniciación N°11190360, Fondecyt Postdoctorado N°3230678, and Fondecyt
Regular N°1210526. The companies Garces Fruit and CER Analytics provided
access to the \emph{in-situ} soil moisture data used in this work.
